---
title: "Interactive Shiny Studio — Complete User Guide"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Interactive Shiny Studio — Complete User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  eval       = FALSE
)
```

## Overview

The **scSimEval Shiny Studio** is a fully self-contained, browser-based graphical interface that
gives researchers access to all 62 evaluation measures and every visualization in the package —
without writing a single line of R code.

Launch it with:

```r
library(scSimEval)
launch_scSimEval_app()          # opens at http://127.0.0.1:<port>
```

> **File size limit:** The app accepts uploads up to **500 MB** per file, which handles most
> real-world single-cell datasets.

The studio is organized into **six primary navigation tabs** in the top bar, plus an
**eight-panel Visualizations module** inside Tab 4. The sections below document every tab,
every control, and all export options in the order they appear in the application.

---

## Tab 1 — Home

The landing page gives a quick orientation and provides **one-click jump buttons** to any part
of the workflow:

| Button | Destination |
|--------|-------------|
| **1. Data Upload & Evaluation** | Data Hub tab |
| **2. Comparative Bubble Matrix** | Bubble Matrix tab |
| **3. Diagnostic Visualizations** | Visualizations tab |
| **4. Download Results** | Download Results tab |
| **5. Help & Manual** | Help & Getting Started tab |
| **6. Team & Contact** | Team & Contact tab |

The four **KPI cards** below the hero section display: 62 curated evaluation measures,
8 canonical categories, 3 supported modalities, and the [0, 1] standardized fidelity scale.

---

## Tab 2 — Data Hub

Data ingestion and evaluation control centre. Choose one of **four evaluation modes**
from the radio button panel on the left sidebar.

### Mode 1 — Explore Demo Benchmark (6 Simulators)

Click **Load Demo Benchmark (6 Simulators)** to instantly populate the app with pre-computed
results for **Splatter**, **scDesign3**, **SCRIP**, **SymSim**, **dyngen**, and **simATAC**
across all 62 measures. No file uploads required.

### Mode 2 — Single-Cell Evaluation (scRNA-seq / scATAC-seq)

Evaluate one or more unimodal count-matrix simulators.

| Step | Control | Notes |
|------|---------|-------|
| **1** | **Reference Count Matrix** | Real empirical cells. Formats: `.rds`, `.csv`, `.tsv`, `.txt`, `SingleCellExperiment`, `Seurat`. |
| **2** | **Simulated Count Matrices** (multi-select) | One file per simulator. Same formats. |
| **3** | **Simulator Name / Runtime (s) / Memory (MiB)** | Enter for each uploaded simulator. |
| **4** | **Cell Type Labels** (optional) | `.rds` (factor / vector / data.frame) or `.csv`/`.tsv`. |
| **4** | **Batch Annotations** (optional) | Same formats as cell types. |
| — | **Append to current benchmark** | Adds new results alongside existing ones. |
| — | **Evaluate Single-Cell Simulators** | Triggers the full 50+ metric pipeline. |

### Mode 3 — Multiomics Evaluation (scRNA-seq + scATAC-seq)

| Subtype | Description |
|---------|-------------|
| **Paired Co-assay** | Same individual cells in both modalities (10x Multiome, SHARE-seq). All 62 metrics including FOSCTTM and Match@1. |
| **Unpaired Profiling** | Separate cells from same tissue. Population-level cross-modal metrics only; cell-pairing metrics safely skipped. |

| Step | Control | Notes |
|------|---------|-------|
| **1** | **Reference RNA Count Matrix** | Genes × cells. |
| **1** | **Reference ATAC Count Matrix** | Peaks × cells. |
| **2** | **Number of Multiomics Simulators** | 1 – 5. Generates matching upload fields. |
| **2** | **Per-Simulator Fields** | Simulated RNA matrix, simulated ATAC matrix, name, runtime (s), memory (MiB). |
| **3** | **Cell Type Labels / Batch** (optional) | Applied to RNA modality. |
| — | **Evaluate Multiomics Simulators** | Runs the full 62-metric multiomics pipeline. |

### Mode 4 — Upload Saved Results (.rds / .csv)

Load output previously saved from `evaluate_simulation_accuracy()`,
`evaluate_multiomics_accuracy()`, or `evaluate_multiple_datasets()`:

```r
bench <- evaluate_simulation_accuracy(ref, sim, method_name = "MyTool")
saveRDS(bench, "my_benchmark.rds")
```

Click **Load Saved File** to restore the entire benchmark session.

### Main Panel — Active Benchmark & Dataset Summary

| Section | Content |
|---------|---------|
| **Status Banner** | Number of simulators loaded and total metrics computed |
| **Dataset Properties Table** | Per-matrix: Cells (N), Features (P), Sparsity (%), Cell Types, Batches, Median Library Size, Median Detected Features, Mean Expression |
| **Download Dataset Summary** | Export properties table as CSV |
| **Benchmark Preview Table** | Searchable / sortable DT table of all raw and normalized metric scores |

---

## Tab 3 — Comparative Bubble Matrix

The flagship visualization: every simulator × all 62 metrics in one colour-coded matrix.
Bubble **diameter** is proportional to the normalized fidelity score.
Scores ≥ 0.96 are rendered as bold square glyphs.
Simulators are rank-ordered top-to-bottom by composite average score.

The main panel organizes content into two dedicated subtabs:
1. **Comparative Bubble Matrix Figure**: Displays the full 62-metric matrix figure with smooth horizontal panning at publication resolution.
2. **Fidelity Leaderboard**: Positioned right after the figure subtab, displaying the complete rankings table (Overall Rank, Method, Average Fidelity %, Fidelity Score) with an instant CSV export button.

### Sidebar Controls

| Control | Options | Default |
|---------|---------|---------|
| **Category Filter** | All 62 · (I) Distributional Properties · (II) Correlations & Zero-Inflation · (III) Cellular Structure & Concordance · (IV) Batch Effects & Confounder Mixing · (V) Biological Signal & Downstream Fidelity · (VI) Trajectory & Lineage Dynamics · (VII) Cross-Modal Coupling & Modularity · (VIII) Computational Scalability | All |
| **Method Selector** | Dynamic checkboxes per simulator | All selected |
| **Matrix Display Width (px)** | 1 200 – 3 200 (step 50) | 2 200 |
| **Matrix Display Height (px)** | 450 – 1 100 (step 20) | 680 |

> **Tip:** Default 2 200 × 680 px matches the publication format (23 : 7 aspect ratio). Use the
> horizontal scrollbar to inspect all 62 measures without label squishing.

### Exports

| Button | Output |
|--------|--------|
| **Download JPEG (600 DPI)** | Publication-grade `.jpeg` |
| **Download Vector PDF** | Scalable `.pdf` |
| **Download Leaderboard (CSV)** | Tidy CSV table of accurate method rankings and fidelity scores |

---

## Tab 4 — Visualizations

Eight diagnostic sub-panels accessed via the **pill navigation bar**. Every sub-panel
exposes individual 600 DPI JPEG and vector PDF export buttons.

### Sub-panel 1 — Evaluation Summary

`plot_evaluation_summary()` — rank-ordered horizontal bar chart.

| Control | Description |
|---------|-------------|
| **Show Score Labels** | Toggle numerical scores on each bar |
| **Normalize Scores [0, 1]** | Standardized vs. raw display |

### Sub-panel 2 — Distribution QC

`plot_distribution_qc()` — 14-panel grid comparing expression densities,
library sizes, and zero-inflation curves between reference and simulated cells.

| Control | Options |
|---------|---------|
| **QC Layout** | Comprehensive (14 panels) · Density Curves Only |

### Sub-panel 3 — Scalability Benchmark

`plot_scalability_benchmark()` — runtime and memory dashboards with Pareto efficiency frontiers.

| Control | Options |
|---------|---------|
| **Scalability View** | 4-Panel Comprehensive · Runtime Only · Peak RAM Only · Runtime vs Memory Trade-Off · Resource Cost Footprint |

### Sub-panel 4 — Metric Plots

`plot_metric_boxplots()` / `plot_category_metric_bars()` / `plot_individual_metric_bar()` — two view modes.

**Individual Metric mode:**

| Control | Description |
|---------|-------------|
| **1. Choose Category** | Filters the metric dropdown |
| **2. Choose Metric** | Dynamic; populated by selected category |
| **3. Score Type** | Normalized [0, 1] · Raw Metric Value |

**Category Group mode:**

| Control | Description |
|---------|-------------|
| **Choose Category** | All 8 combined (displays 8-category distribution boxplots via `plot_metric_boxplots()`), or an individual category (displays faceted barplots across all category metrics via `plot_category_metric_bars()`) |
| **Score Type** | Normalized [0, 1] · Raw Value |

### Sub-panel 5 — Metric Heatmap

`plot_metric_heatmap()` — method × metric grid; exact raw scores in bold text
with direction-aware fill colours.

| Control | Options | Default |
|---------|---------|---------|
| **Category Filter** | All 8 combined · individual category | All |
| **Heatmap Height (px)** | 400 – 1 500 (step 50) | 1 100 |
| **Heatmap Width (px)** | 600 – 1 300 (step 20) | 880 |

### Sub-panel 6 — PCA Ordination

`plot_metric_pca()` — simulators projected into performance space with discriminating
metric vector loadings.

| Control | Options | Default |
|---------|---------|---------|
| **Category Filter** | All Combined · (I) Dist. Props · (II) Corr. & Zero-Inflation · (III) Cellular Structure · (IV) Batch Effects · (V) Bio. Signal · (VII) Cross-Modal | All |
| **Panel View** | Both (Biplot + Loadings) · Simulators Only · Loadings Only | Both |
| **Top Metrics (Loadings)** | 4 – 50 (step 1) | 14 |
| **PCA Height (px)** | 500 – 1 600 (step 50) | 1 100 |
| **PCA Width (px)** | 650 – 1 400 (step 25) | 950 |

> When **All Categories Combined** is selected, only top-ranked metrics by loading magnitude
> [√(PC1² + PC2²)] are shown to prevent visual clutter.

### Sub-panel 7 — MDS Metric Space

`plot_metric_mds()` — Multi-Dimensional Scaling ordination.

| Control | Options | Default |
|---------|---------|---------|
| **Category Filter** | Same six options as PCA | All |
| **MDS Target** | By Simulators · By Metric Summaries | By Simulators |
| **MDS Height (px)** | 400 – 1 300 (step 20) | 620 |
| **MDS Width (px)** | 600 – 1 400 (step 20) | 900 |

### Sub-panel 8 — Cell Embeddings (t-SNE & UMAP)

`compute_dataset_embeddings()` + `plot_dataset_embeddings()`.

**Reduction & Layout:**

| Control | Options | Default |
|---------|---------|---------|
| **Reduction Technique** | UMAP · t-SNE · PCA | UMAP |
| **Comparison Layout** | Faceted Grid (All Simulators) · Side-by-Side (vs Single Simulator) | Faceted Grid |
| **Select Simulator** (side-by-side only) | Dynamic dropdown | — |

**Colour & Simulator:**

| Control | Options | Default |
|---------|---------|---------|
| **Simulator Selector** | Dynamic checkboxes | All |
| **Color Cells By** | Biological Cell Type / Group · Unsupervised Cluster · Sequencing Depth (Library Size) · Technical Batch · Dataset Source | Cell Type |

**Fine-Tuning:**

| Control | Range | Default |
|---------|-------|---------|
| **Number of PCs** | 5 – 50 (step 5) | 20 |
| **t-SNE Perplexity** (t-SNE only) | 5 – 50 (step 5) | 15 |
| **UMAP Neighbors** (UMAP only) | 5 – 50 (step 5) | 15 |
| **Point Size** | 0.2 – 3.0 (step 0.1) | 1.0 |
| **Alpha (transparency)** | 0.2 – 1.0 (step 0.05) | 0.8 |

**Quantitative Embedding Quality Table** (below the plot):

| Column | Description |
|--------|-------------|
| Dataset | Reference or simulator name |
| Cells | Number of cells |
| Mean Silhouette | Average silhouette width of cell-type clusters |
| ARI (Cluster Fidelity) | ARI between k-means clusters and known labels |
| Mean Library Size | Mean total UMI / peak count per cell |
| Mean Detected Features | Mean genes / peaks detected per cell |
| Library Size Diff (%) | Absolute % difference vs. reference mean |

---

## Tab 5 — Download Results

### 1. All-in-One Benchmark Archive (.zip)

The comprehensive benchmark archive packages all quantitative results, tabular summaries, compiled reports, and figures into an organized directory structure:

- **Metrics in Multiple Formats (`metrics/` and root):**
  - **Excel Workbook (`.xlsx`)** — 4 sheets: All Benchmark Metrics, Dataset Properties, Method Rankings, Scalability Metrics
  - **Master CSV Table (`.csv`)** — All 62 metrics in tidy long format
  - **Master TXT Table (`.txt`)** — Tab-delimited text export of the complete metric evaluation matrix
  - **Method Rankings Leaderboard (`.csv` and `.txt`)** — Overall scores, ranks, and category-level rankings
  - **Dataset Properties Summary (`.csv` and `.txt`)** — Cell counts, feature counts, sparsity, library size, and detected features
  - **Native R Object (`.rds`)** — Serialized R list containing all benchmark objects for reproducible downstream analysis and custom plotting
- **Compiled PDF Report (`scSimEval_all_plots_report.pdf`):** Multi-page publication-ready vector report containing all 9 benchmark figures
- **Grouped Figures (`figures/grouped/`):** Both 600 DPI publication-grade JPEGs and vectorized PDFs for all overview diagnostic figures:
  1. Comparative Bubble Matrix (`01_bubble_matrix_comparative`)
  2. Overall Evaluation Summary (`02_evaluation_summary_grouped`)
  3. Scalability Benchmark (`03_scalability_benchmark_grouped`)
  4. Metric Boxplots by Category (`04_metric_boxplots_by_category`)
  5. Performance Heatmap (`05_metric_performance_heatmap`)
  6. PCA Simulator Ordination (`06_simulator_pca_ordination`)
  7. MDS Ordination (`07_simulator_mds_ordination`)
  8. Distribution QC Curves (`08_comparative_distribution_qc`)
  9. Cell Embeddings UMAP Grid (`09_cell_embeddings_umap_grid`)
  10. Cell Embeddings t-SNE Grid (`10_cell_embeddings_tsne_grid`)
  11. Cell Embeddings PCA Grid (`11_cell_embeddings_pca_grid`)
- **Individual Figures (`figures/individual/`):**
  - High-resolution bar plots for each individual metric (`metric_bar_<MetricName>.jpeg` & `.pdf`)
  - Simulator-specific 1-to-1 comparison cell embeddings against empirical reference across all reductions (`cell_embeddings_compare_<reduction>_<SimulatorName>.jpeg` & `.pdf` for UMAP, t-SNE, and PCA)

Click **Download Complete Results (.zip)**.

### 2. Spreadsheets & Data Files

| Button | Format | Description |
|--------|--------|-------------|
| Download Excel File (.xlsx) | Multi-sheet workbook (`.xlsx`) | Full benchmark scores, rankings, dataset metadata, and scalability |
| Download CSV Table (.csv) | Flat tidy table (`.csv`) | Comma-separated values for Python (pandas), Excel, or R |
| Download TXT Table (.txt) | Tab-delimited table (`.txt`) | Clean TSV/TXT export for command-line workflows and bioinformatics tools |
| Download R Data File (.rds) | Native R serialized object (`.rds`) | Ready for `readRDS()` in downstream R pipelines |

### 3. Complete Multi-Page PDF Report

All figures compiled into a single high-quality vector PDF report:

1. Comparative Bubble Matrix
2. Overall Evaluation Summary
3. Scalability Benchmark
4. Metric Boxplots by Category
5. Performance Heatmap
6. PCA Simulator Ordination
7. MDS Ordination
8. Distribution QC Curves
9. Cell Embeddings Comparison Grids

Click **Download All Figures (.pdf)**.

### 4. Export Individual Figures (High-Res 600 DPI JPEG & Vector PDF)

Download any single visualization on demand with dedicated format controls:

| Control | Description |
|---------|-------------|
| **Choose Figure to Export** | Dropdown spanning all 14 benchmark visual types: Comparative Bubble Matrix, Overall Evaluation Summary, Scalability Benchmark, Metric Boxplots, Performance Heatmap, PCA Ordination, MDS Metric Space, Distribution QC, UMAP Comparison Grid, t-SNE Comparison Grid, PCA Comparison Grid, or 1-to-1 side-by-side comparison against reference (UMAP, t-SNE, PCA) |
| **Select Simulator (for 1-to-1 compare)** | Dynamically populated dropdown to pick the specific simulator to benchmark against the empirical reference |
| **Download Vector PDF (.pdf)** | Vectorized, infinitely scalable PDF styled to optimal aspect ratio |
| **Download 600 DPI JPEG (.jpeg)** | Ultra-high-resolution raster JPEG ready for high-impact journal submission |

### Interactive Benchmark Data Table

A full `DT` table with **per-column search boxes** lets you filter by Method Name, Category,
or Metric before exporting.

---

## Tab 6 — Help & Getting Started

An in-app reference guide structured into seven sections:

1. **Workflow Architecture** — 4-step pipeline diagram
2. **Eight Evaluation Categories** — tabular reference for all 62 metrics
3. **Score Normalization Pipeline** — direction inversion → min-max scaling diagrams
4. **Studio Navigation Guide** — accordion-style guide to each tab
5. **Accepted Input Formats** — matrix types, annotation vectors, saved files
6. **Frequently Asked Questions**
7. **Team, Contact & Citation**

The tab also prominently links to the full online documentation at
<https://kabilanbio.github.io/scSimEval>.

---

## R API Cross-Reference

Every Shiny control maps directly to a documented R function:

| Shiny Panel | R Function |
|-------------|-----------|
| Data Hub → Single-Cell | `evaluate_simulation_accuracy()` |
| Data Hub → Multiomics | `evaluate_multiomics_accuracy()` |
| Data Hub → Multiple Datasets | `evaluate_multiple_datasets()` |
| Bubble Matrix | `plot_benchmark_bubble_matrix()` |
| Evaluation Summary | `plot_evaluation_summary()` |
| Distribution QC | `plot_distribution_qc()` |
| Scalability Benchmark | `plot_scalability_benchmark()` |
| Metric Plots (all categories) | `plot_metric_boxplots()` |
| Metric Plots (by category) | `plot_category_metric_bars()` |
| Metric Plots (individual) | `plot_individual_metric_bar()` |
| Metric Heatmap | `plot_metric_heatmap()` |
| PCA Ordination | `plot_metric_pca()` |
| MDS Metric Space | `plot_metric_mds()` |
| Cell Embeddings | `compute_dataset_embeddings()` + `plot_dataset_embeddings()` |
| Embedding Quality Metrics | `compute_embedding_quality_metrics()` |

For complete parameter descriptions see the [Reference](../reference/index.html) index.

---

```{r session-info, eval=TRUE, echo=FALSE}
sessionInfo()
```
